---
title: "Working with Dates and Regex"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lubridate)
library(dplyr)
library(ggplot2)
```

## Purpose
Explore Biketown data looking at the mean miles traveled per week during the summer of 2018.

```{r pull_data}
source("code/fetch_biketown.R")
get_data(start = "06/2018", end = "09/2018")

setwd("/Users/lslawsk/Documents/RTraining/datascience2019/data/biketown/")

folder <- "/Users/lslawsk/Documents/RTraining/datascience2019/data/biketown/"
filenames <- list.files(path = folder, pattern = ".csv", all.files = FALSE,
                        full.names = FALSE,
                        recursive = FALSE, ignore.case = FALSE)

read_csv_filename <- function(filenames){
  ret <- read.csv(filenames, stringsAsFactors = F,
                  strip.white = T, na.strings = "")
  ret$Source <- filenames
  ret
}
bike_raw <- plyr::ldply(filenames, read_csv_filename)

#### if doesn't work then just call in one file at a time
# file1 <- read.csv("/Users/lslawsk/Documents/RTraining/datascience2019/data/biketown/2018_06.csv",
# stringsAsFactors = F)
# bike_raw2 <- bindrows(file1)

```

```{r data_wrangling}
str(bike_raw)

bike_df1 <- bike_raw %>% 
  mutate(start.datetime = paste(StartDate, StartTime, sep = " "),
         end.datetime = paste(EndDate, EndTime, sep = " "))
str(bike_df1)

# convert chr format of last 2 columns to date-time format
# !!!!make sure to include timezone!!!!
# check acceptable timezone values
OlsonNames()
bike_df1$start.datetime <- mdy_hm(bike_df1$start.datetime,
                                  tz = "America/Los_Angeles")
bike_df1$end.datetime <- mdy_hm(bike_df1$end.datetime,
                                tz = "America/Los_Angeles")
str(bike_df1)

# convert duration into usable format
bike_df1$Duration <- hms(bike_df1$Duration)
# error thrown after convert indicating NAs; check raw data
sum(is.na(bike_df1$Duration))
```

'lubridate' has three synonymous functions but they define very different actions:
1. duration(): time spans are in units of seconds, and there is no start date involved.
2. interval(): also measured in units of seconds, and provides the difference between start and end.
3. period)(): measures time span in units larger than seconds; this is handy when accounting for daylight saving time and leap years.

``` {r duration_calc}
# calculate interval
bike_df1$interval <- interval(bike_df1$start.datetime, bike_df1$end.datetime)
head(bike_df1$interval)

# calculate duration. Must convert to interval first
bike_df1$duration.all <- as.duration(bike_df1$interval)
head(bike_df1$duration.all)

# calculate period
bike_df1$period <- as.period(bike_df1$duration.all)
head(bike_df1$period)

# floor_date() is helpful for aggregating
bike_wkagg <- bike_df1 %>% 
  mutate(week.datetime = floor_date(start.datetime, unit = "week"))
head(bike_wkagg$week.datetime)

# calculate mean distance traveled per week
mean_wkdist <- bike_wkagg %>% 
  filter(Source %in% c("2018_06.csv","2018_07.csv","2018_08.csv","2018_09.csv")) %>% 
  group_by(week.datetime) %>% 
  summarize(wk_mean_dist = mean(Distance_Miles))

plot1 <- ggplot(mean_wkdist, aes(as.Date(week.datetime), wk_mean_dist)) +
  geom_bar(stat = "identity", fill = "orange") +
  labs(title = "Average Biketown Ride Distance by Week",
       subtitle = "July-Sept 2018",
       caption = "source: Biketown") +
  xlab("Week Of") + ylab("Average Miles") +
  scale_x_date(date_breaks = "1 week") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_bw()
plot1

```
